Advanced Reactor Fuel Cycle (ARFC) Molten Salt Reactor Design

### 2016-10-06 MSR-design.tex
include packages ordering is always complicated by hyperref
most style guides prefer the reference within the sentence (e.g. before the period).

\documentclass{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex plus 0.5ex minus 0.2ex}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}
\usepackage{textcomp}
\usepackage{placeins}
\usepackage[T1]{fontenc}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage{caption}
\usepackage[export]{adjustbox}
\graphicspath{{./figures/}}

\usepackage[acronym,toc]{glossaries}
\include{acros}
\makeglossaries

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{salt_properties.png}
  \caption{Composition and properties of fuel, flush, and coolant
    salts. \cite{robertson_msre_1965}}
  \label{fig:salt_properties}
\end{figure}

@@

A cutaway diagram of the \gls{MSRE} reactor is shown in \cref{fig:MSRE_reactor}. Fuel
enters through the inlet and is distributed evenly along the outside of the
reactor vessel in a small annular region of 1 inch
width. \cite{robertson_msre_1965} The fluid fuel then flows turbulently
downward, cooling the reactor vessel to within 5\textdegree F of the entering
bulk salt temperature (1175 \textdegree F). \cite{robertson_msre_1965} Some \gls{MSRE}
modeling papers (\cite{kophazi_development_2009}) refer to this down-flowing
region as the \textit{downcomer}. The fuel salt is collected in the lower plenum
(lower \textit{space}) of the reactor where straightening vanes remove the
salt's rotational motion. The salt then flows upward through approximately 1140
0.4-inch by 1.2-inch fuel channels that are formed by machining grooves in the
sides of the graphite moderator columns \cite{robertson_msre_1965}; these are the
regions where fission occurs. The graphite columns themselves have 2-inch by
2-inch square cross sections. Nominal core volume is 90 ft$^3$, with 70 ft$^3$
occupied by graphite and 20 ft$^3$ occupied by fuel. \cite{robertson_msre_1965}
After flowing through the core channels, the fuel collects in the upper plenum
before exiting through the outlet.

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{MSRE_reactor_vessel.png}
  \caption{\gls{MSRE} reactor vessel. \cite{robertson_msre_1965}}
  \label{fig:MSRE_reactor}
\end{figure}

@@

The \gls{MSRE} has a negative temperature coefficient of $6.4 - 9.9 \times10^{-5}$
($\Delta$k/k)/\textdegree F depending on the choice of
fuel. \cite{robertson_msre_1965} The \gls{MSRE}'s three control rods have a combined
worth of 5.6 - 7.6 \% $\Delta$k/k, again depending on the
fuel. \cite{robertson_msre_1965}

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{reactivity-losses.png}
  \caption{Losses in reactivity associated with different
    phenomena. \cite{robertson_msre_1965}}
  \label{fig:reactivity_losses}
\end{figure}

@@

Kophazi developed a 3D time-dependent simulation that was validated against \gls{MSRE}
data. \cite{kophazi_development_2009} The model consists of a conventional
neutronics model extended to deal with the precursor drift that is unique to
MSRs and the core's thermal behavior. The thermal model includes the true
geometry of the fuel channels as well as the moderator connecting them. Salt
flow is assumed to be only in the reactor's axial direction, which reduces the
number of non-linear variables that have to be solved for.

@@

with the last term representing the effect of fuel advection. In
\cite{kophazi_development_2009}, \cref{eq:neutrons} is solved only in the
reactor core, whereas \cref{eq:precursors} is solved both in the core and in the
\gls{MSRE} external primary circuit. The external loop is modeled as a single pipe
with a volume equal to the total volume of all the external components and
pipelines of the primary circuitry. \cite{kophazi_development_2009} It is not
entirely clear how pipe cross sectional areas are handled, which would determine
the transit times of individual pieces of fluid and subsequently the
distribution of delayed neutron production in the external primary circuit. That
is a piece that would likely need to be included in detailed ARFC accident
models.

@@

The most detailed report on \gls{ORNL}'s molten salt breeder reactor is the technical
report by Robertson. \cite{robertson_conceptual_1971} A nice overview of the
MSBR geometry is shown in \cref{fig:vertical,fig:horizontal,fig:zoom_horiz}.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{Cammi_simplification_1.png}
  \caption{Cammi's simplification of MSBR's geometry. \cite{cammi_multi-physics_2011}}
  \label{fig:simlification}
\end{figure}

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{Cammi_simplification_2.png}
  \caption{(b) Fluid flow and heat transfer boundary conditions; (c) neutron flux and DNP boundary conditions. \cite{cammi_multi-physics_2011}}
  \label{fig:simlification-2}
\end{figure}

>>>>

\documentclass{article}
\setlength{\parindent}{0pt}
\setlength{\parskip}{2ex plus 0.5ex minus 0.2ex}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{placeins}
\usepackage[T1]{fontenc}
\usepackage{gensymb}
\usepackage[utf8]{inputenc}
\usepackage{caption}
\usepackage[export]{adjustbox}
\graphicspath{{./figures/}}
% hyperref usually has to go last
\usepackage[hidelinks]{hyperref}
% but glossaries behaves best if after hyperref
\usepackage[acronym,toc]{glossaries}
\include{acros}
\makeglossaries
% cleveref only behaves if after hyperref & glossaries
\usepackage{cleveref}

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{salt_properties.png}
  \caption{Composition and properties of fuel, flush, and coolant
    salts \cite{robertson_msre_1965}.}
  \label{fig:salt_properties}
\end{figure}

@@

A cutaway diagram of the \gls{MSRE} reactor is shown in \cref{fig:MSRE_reactor}. Fuel
enters through the inlet and is distributed evenly along the outside of the
reactor vessel in a small annular region of 1 inch
width \cite{robertson_msre_1965}. The fluid fuel then flows turbulently
downward, cooling the reactor vessel to within 5\textdegree F of the entering
bulk salt temperature (1175 \textdegree F) \cite{robertson_msre_1965}. Some \gls{MSRE}
modeling papers (\cite{kophazi_development_2009}) refer to this down-flowing
region as the \textit{downcomer}. The fuel salt is collected in the lower plenum
(lower \textit{space}) of the reactor where straightening vanes remove the
salt's rotational motion. The salt then flows upward through approximately 1140
0.4-inch by 1.2-inch fuel channels that are formed by machining grooves in the
sides of the graphite moderator columns \cite{robertson_msre_1965}; these are the
regions where fission occurs. The graphite columns themselves have 2-inch by
2-inch square cross sections. Nominal core volume is 90 ft$^3$, with 70 ft$^3$
occupied by graphite and 20 ft$^3$ occupied by fuel \cite{robertson_msre_1965}.
After flowing through the core channels, the fuel collects in the upper plenum
before exiting through the outlet.

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{MSRE_reactor_vessel.png}
  \caption{\gls{MSRE} reactor vessel \cite{robertson_msre_1965}.}
  \label{fig:MSRE_reactor}
\end{figure}

@@

The \gls{MSRE} has a negative temperature coefficient of $6.4 - 9.9 \times10^{-5}$
($\Delta$k/k)/\textdegree F depending on the choice of
fuel \cite{robertson_msre_1965}. The \gls{MSRE}'s three control rods have a combined
worth of 5.6 - 7.6 \% $\Delta$k/k, again depending on the
fuel \cite{robertson_msre_1965}.

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{reactivity-losses.png}
  \caption{Losses in reactivity associated with different
    phenomena \cite{robertson_msre_1965}.}
  \label{fig:reactivity_losses}
\end{figure}

@@

Kophazi developed a 3D time-dependent simulation that was validated against \gls{MSRE}
data \cite{kophazi_development_2009}. The model consists of a conventional
neutronics model extended to deal with the precursor drift that is unique to
MSRs and the core's thermal behavior. The thermal model includes the true
geometry of the fuel channels as well as the moderator connecting them. Salt
flow is assumed to be only in the reactor's axial direction, which reduces the
number of non-linear variables that have to be solved for.

@@

with the last term representing the effect of fuel advection. In
\cite{kophazi_development_2009}, \cref{eq:neutrons} is solved only in the
reactor core, whereas \cref{eq:precursors} is solved both in the core and in the
\gls{MSRE} external primary circuit. The external loop is modeled as a single pipe
with a volume equal to the total volume of all the external components and
pipelines of the primary circuitry \cite{kophazi_development_2009}. It is not
entirely clear how pipe cross sectional areas are handled, which would determine
the transit times of individual pieces of fluid and subsequently the
distribution of delayed neutron production in the external primary circuit. That
is a piece that would likely need to be included in detailed ARFC accident
models.

@@

The most detailed report on \gls{ORNL}'s molten salt breeder reactor is the technical
report by Robertson \cite{robertson_conceptual_1971}. A nice overview of the
MSBR geometry is shown in \cref{fig:vertical,fig:horizontal,fig:zoom_horiz}.

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{Cammi_simplification_1.png}
  \caption{Cammi's simplification of MSBR's geometry 
        \cite{cammi_multi-physics_2011}.}
  \label{fig:simlification}
\end{figure}

@@

\begin{figure}[htpb]
  \centering
  \includegraphics[max height=.5\textheight,max width=\textwidth,keepaspectratio]{Cammi_simplification_2.png}
  \caption{(b) Fluid flow and heat transfer boundary conditions; (c) neutron 
        flux and DNP boundary conditions \cite{cammi_multi-physics_2011}.}
  \label{fig:simlification-2}
\end{figure}
